---
title: "week6"
format: html
editor: visual
---

## `Browser()`

```{r}
library(tidyverse)
library(palmerpenguins)

traits_r2 <- function(trait1, trait2) {
  #browser()  #Put browser here!! 
  summary(lm(trait1 ~ trait2))$r.squared
}

penguins %>% 
  summarize(flipper_bill_r2 = traits_r2(flipper_length_mm, bill_length_mm))
```

**Q1:** How would you describe the error? I\'m not asking you describe the *cause* of the error yet. Describe how the output is different than what you would expect.

*I am expecting a tibble with content (i.e. a column), but there are no columns.*

**Q2:** Where would you add `browser()` to explore the cause of the error?

*Put it before the line "summary(lm(trait1 \~ trait2))\$rsquared" in the function.*

**Q3:** Does the body of `traits_r2` use list or atomic vector indexing? Does it use indexing by position, logic, or name?

*It uses list indexing (named list).*

**Q4:** What\'s the cause of the error? How would you fix it?

*The error is a typo: "rsquared" should be "r.squared".*

## **Adding `group_by()`**

```{r}
# Pipeline 1
penguins %>% 
  group_by(species) %>% 
  summarize(flipper_bill_r2 = traits_r2(flipper_length_mm, bill_length_mm))

# Pipeline 2
penguins %>% 
  group_by(species, island) %>% 
  summarize(flipper_bill_r2 = traits_r2(flipper_length_mm, bill_length_mm),
            .groups = "drop")
```

**Q5:** How many times does Pipeline 1 call `traits_r2()`? How about Pipeline 2?

Pipeline 1 calls `traits_r2()` 3 times (3 species). Pipeline 2 calls `traits_r2()` 5 times.

**Q6:** Create Pipeline 3 that additionally groups by sex. How many times does Pipeline 3 call `traits_r2()`?

```{r}
# Pipeline 3
penguins %>% 
  group_by(species, island, sex) %>% 
  summarize(flipper_bill_r2 = traits_r2(flipper_length_mm, bill_length_mm),
            .groups = "drop") 
```

Pipeline 3 calls `traits_r2()` 13 times.

## **Error in `group_by()`-`summarize()`**

```{r}
set.seed(12345)
penguins2 <- penguins %>% 
  drop_na(sex) %>% 
  sample_n(25)
penguins2[7, 3:6] <- NA


traits_r2 <- function(trait1, trait2) {
#  browser()  #Put browser here!! 
  summary(lm(trait1 ~ trait2))$r.squared
}

penguins2 %>% 
  drop_na() %>% 
  group_by(species, island, sex) %>% 
  summarize(flipper_bill_r2 = traits_r2(flipper_length_mm, bill_length_mm),
            .groups = "drop") 
```

# **Visualizing data with `ggplot()`**

```{r}
library(palmerpenguins)
library(tidyverse)

ggplot(data = penguins, aes(x = body_mass_g, y = flipper_length_mm)) +
  geom_point()
```

**Q10:** Change the aesthetics of the plot to show `bill_depth_mm` on the y-axis.

```{r}
ggplot(data = penguins, aes(x = body_mass_g, y = bill_depth_mm)) +
  geom_point()
```

**Q11:** Add an aesthetic for color, so points will be color-coded by species.

```{r}
ggplot(data = penguins, 
       aes(x = body_mass_g, y = bill_depth_mm,
           color = species)) +
  geom_point()
```

## **Customizing scales**

```{r}
ggplot(data = penguins) +
  # You can also define aesthetics directly inside the geometries
  geom_point(aes(x = body_mass_g, 
                 y = flipper_length_mm,
                 color = species)) +
  # x aesthetic, continuous scale
  scale_x_continuous(
    # change the axis name
    name = "Body mass (g)",
    # change the limits
    limits = c(2000, 8000)
  ) +
  # color aesthetic, manual scale
  scale_color_manual(
    # set the values for the colors
    values = c(Adelie = "cornflowerblue",
               Chinstrap = "firebrick",
               Gentoo = "darkorchid")
  )
```

**\
Q12:** What function would you use to customize the y-axis? Use that function to expand the y-axis limits to include 150 and 250.

```{r}
ggplot(data = penguins) +
  # You can also define aesthetics directly inside the geometries
  geom_point(aes(x = body_mass_g, 
                 y = flipper_length_mm,
                 color = species)) +
  # x aesthetic, continuous scale
  scale_x_continuous(
    # change the axis name
    name = "Body mass (g)",
    # change the limits
    limits = c(2000, 8000)
  ) +
  # color aesthetic, manual scale
  scale_color_manual(
    # set the values for the colors
    values = c(Adelie = "cornflowerblue",
               Chinstrap = "firebrick",
               Gentoo = "darkorchid")
  )+
  #This is the new section
  scale_y_continuous(limits = c(150, 250))
```

**Q13:** Look up the help for `scale_color_brewer()`. Change the color scale to use the \"Dark2\" color palette.

```{r}
ggplot(data = penguins) +
  # You can also define aesthetics directly inside the geometries
  geom_point(aes(x = body_mass_g, 
                 y = flipper_length_mm,
                 color = species)) +
  # x aesthetic, continuous scale
  scale_x_continuous(
    # change the axis name
    name = "Body mass (g)",
    # change the limits
    limits = c(2000, 8000)
  ) +
  # This is the new section
  scale_color_brewer(palette = "Dark2")
```

**Q14:** Use `geom_histogram()` to visualize the distribution of body masses.

```{r}
ggplot(penguins, aes(x=body_mass_g))+
  geom_histogram()
```

**Q15:** Use `geom_boxplot()` to create boxplots of the flipper lengths by species. It should look like the figure below. Hint: what aesthetics will you use?

```{r}
ggplot(penguins, aes(x=species, y=flipper_length_mm))+
  geom_boxplot()
```

**Q16:** `geom_pointrange()` is like `geom_point()` with the addition of error bars. Like `geom_point()`, `geom_pointrange()` takes `x` and `y` aesthetics, but it also takes `xmin`, `xmax`, `ymin` and `ymax` aesthetics. Create a figure that shows the mean ± sd of each penguin species\' body mass and flipper lengths using points and error bars; it should look like the figure below. First, create a summary data frame that calculates the mean and standard deviation of the two variables you need by each species. Then use that summary data frame to create your figure. Hint: `geom_pointrange()` takes `xmin` ***or*** `xmax`, same for `ymin` ***or*** `ymax`. So you\'ll need to call `geom_pointrange()` twice and specify the aesthetics within the geometry.

```{r}
plot_df <- penguins %>% 
  drop_na() %>% 
  group_by(species) %>% 
  summarise(body_mass_mean = mean(body_mass_g),
            body_mass_sd = sd(body_mass_g),
            flipper_length_mean = mean(flipper_length_mm),
            flipper_length_sd = sd(flipper_length_mm))

ggplot(plot_df, aes(color = species))+
  geom_pointrange(aes(
    x = body_mass_mean, 
    y = flipper_length_mean,
    xmin = body_mass_mean-body_mass_sd,
    xmax = body_mass_mean+body_mass_sd))+
  geom_pointrange(aes(x = body_mass_mean, 
    y = flipper_length_mean, 
    ymin = flipper_length_mean-flipper_length_sd, 
    ymax = flipper_length_mean+flipper_length_sd))
```
